---
title: "Perturbations dans la commune d'Androy"
author: "Florent Bédécarrats, Stéphanie Carrière"
format: 
  html: 
    toc: true
    output-file: index.html
    embed-resources: true
    standalone: true
execute:
  warning: false
  error: false
editor: visual
editor_options: 
  chunk_output_type: console
bibliography: references.bib
---

## A focus on the Androy Municipality

Located in Fianarantsoa Region. We obtain the municipal delimitation from the Global Adminisative Area database (GADM).

```{r admin_data}
#| fig-cap: "Androy municipality location (Source: GADM)"

library(tidyverse)
library(geodata)
library(sf)
library(terra)
library(mapme.biodiversity)
library(tmap)
library(zoo)

if (file.exists("mada.RDS")) {
  mada <- read_rds("mada.RDS")
  communes_mada <- read_rds("communes_mada.RDS")
} else {
  mada <- gadm("MDG", level=0, path = tempdir()) %>%
    st_as_sf() 
  communes_mada <- gadm("MDG", level=4, path = tempdir()) %>%
    st_as_sf() 
  write_rds(mada, "mada.RDS")
  write_rds(communes_mada, "communes_mada.RDS")
}

androy <- communes_mada %>%
  filter(NAME_4 == "Androy")
  
tmap_mode("view")
tm_shape(androy) +
  tm_borders(col = "red") + 
  tm_basemap("OpenStreetMap")
```

The municipality parially overlaps with the Ranomafana national park.

## Rainfall

We use the CHIRPS data from NASA to estimate 3 day average rainfall on the Androy Municipality. We compute the data using the package mapme.biodiversity [@mapme.biodiversity].

```{r rainfall}
#| fig-cap: "3 day mean of daily rainfall in Androy municipality (Source: CHIRPS)"
# Fetch data
aoi <- init_portfolio(androy, years = 1980:2021) %>%
  get_resources("chirps")

aoi2 <- aoi %>%
  calc_indicators("precipitation_chirps",
                  engine = "exactextract",
                  scales_spi = 3,
                  spi_prev_years = 8)
rainfall_androy <- aoi2[[16]][[1]]

rainfall_androy %>%
  mutate(rainfall_3d_mean = rollmean(absolute, k = 3, fill = NA)) %>%
  ggplot(aes(x = dates, y = rainfall_3d_mean)) +
  geom_line()
```

## Cyclones

We use the Tropical cyclone best track data (IBTrACS) to get history of cyclone trajectories and cyclone attributes [@knapp_international_2010], as well as raw code from [@mapme.protectedareas].

```{r cyclones}
#| fig-cap: "Cyclone trajectory in Madagascar since 1980 and Androy municipality location (Sources: IBTrACS and GADM)"
if (!file.exists("ibtracs/IBTrACS.since1980.list.v04r00.lines.shp")) {
  dir.create("ibtracs")
  download.file(url = "https://www.ncei.noaa.gov/data/international-best-track-archive-for-climate-stewardship-ibtracs/v04r00/access/shapefile/IBTrACS.since1980.list.v04r00.lines.zip", 
                destfile = "ibtracs/ibtracs_lines.zip")
  unzip("ibtracs/ibtracs_lines.zip", exdir = "ibtracs")
}

# column description is given here: https://www.ncei.noaa.gov/sites/default/files/2021-07/IBTrACS_v04_column_documentation.pdf
cyclones <- read_sf("ibtracs/IBTrACS.since1980.list.v04r00.lines.shp")

# Create a buffer 300km around Madagascar
mada_buff <- st_buffer(mada, 300)

cyclones$wind_combined <- cyclones %>%
  select(contains("WIND")) %>%
  select(-WMO_WIND) %>%
  st_drop_geometry() %>%
  rowMeans(., na.rm = T)

cyclones_mada <- cyclones %>%
  st_intersection(mada_buff) %>%
  select(Season = SEASON, Name = NAME, Winds_in_knots = wind_combined)

tm_shape(androy) + 
  tm_borders(col = "red") + 
  tm_shape(cyclones_mada) +
  tm_lines(col = "Season", lwd = "Winds_in_knots", scale = 3, palette = "Blues",
           popup.vars = c("Name", "Season", "Winds in knots" = "Winds_in_knots"),
           legend.format = list(big.mark=""), popup.format = list(big.mark=""))

```

## References
